Maximizing Ink in Symmetric Partial Edge Drawings of $k$-plane Graphs

### 2018-01-09 eurocg/paper.tex

\begin{theorem}
	\label{thm:tree}
	Let $ G $ be a simple graph with $m$ edges and $ \Gamma $ a straight-line drawing of $ G $. If the intersection graph $ C = (V,E) $ of $ G $ is a tree with maximum degree $ k \in \mathbb{N} $, problem \maxsped can be solved in $ O(mk + m \log m)) $ time and space.
\end{theorem}
% \begin{proof}
% 	Since $ C $ is known to be a tree, and hence $ |E| = O(m) $, we can compute the intersection graph and the stub lengths in $ O(m\log m) $ time and space.
% 	With Lemma~\ref{lem:compute_vertex} and Recurrence~\ref{rec:tree} we need one pass through the tree to find the roots maximum in $ O(mk) $ time. Adding the bounds yields the desired runtime of $ O(mk + m \log m)) $.
% %	To solve the \maxsped problem on $ G $ we need to compute $ \max \{T_i(r) \mid 0 \leq i \leq k \} $ for the root $ r \in V $ of $ C $ as seen above. Doing this bottom up requires one pass through $ C $ hence by applying Lemma~\ref{lem:compute_vertex} we need $ O(mk) $ time to compute all $ T_i(u) $ for all $ u \in V $. For every vertex we only need to compute $ O(k) $ values and store two of them, namely $ \sollong(u) $ and $ \solshort(u) $, so we do not need more than $ O(mk) $ memory. Adding the bounds yields the desired runtime of $ O(mk + m \log m)) $.
% %
%
%
%
% 	Retrieving the actual stub length for each edge in $ G $ needed to produce the SPED can be done by backtracking the subsolutions which requires at most $ O(mk) $ time.
% \end{proof}

@@

\subsection{Cactus graphs}
\label{sec:cactus}
We now generalize our previous algorithm to cactus graphs. 
Due to space constraints we provide only a sketch of the arguments and omit further details.
Let the intersection graph $ C = (V,E) $ be a \emph{cactus graph}, i.e. a simple graph in which no edge $ e \in E $ is part of more than one cycle. 
For a cactus graph $ C $ the \emph{block-cut tree} is an arbitrarily rooted tree $ \mathcal T = (\mathcal V, \mathcal E) $, where $ \mathcal V $ contains a node for every cycle and for every cut vertex in $ C $. 
We call a node $ u \in \mathcal V $ a \emph{block node} if it represents an induced cycle on vertices $ V(u) \subseteq V $ and an \emph{articulation node} of the cut vertex $ a(u) = v \in V $ otherwise. 
%For a block node $ u \in \mathcal V $ we define $ \textit{cycle}(u) = V_c$ and f
%For an articulation node $ u \in \mathcal V $ we define $ a(u) = v $ to be the corresponding cut vertex. 
We have an edge $ uv \in \mathcal E $ between a block node $ u $ and an articulation node $ v $ whenever $ a(v) \in V(u) $ and an edge $ uv \in \mathcal E $ between two articulation nodes whenever $ a(u)a(v) \in E$. % is an edge in $ E $.
We re-use the notation $p(u)$ and $c(u)$ to refer to the parent and to the children of a node $u \in \mathcal V$.

%Let $ C = (V,E) $ be a cactus graph and moreover the intersection graph of a graph $ G $ and straight-line drawing $ \Gamma $. Additionally let $ \mathcal T = (\mathcal V, \mathcal E) $ be the block-cut tree of $ C $ 
We further define for each node $ u \in \mathcal V $ two indices $ i_p(u) \le i_q(u) $. %, w.l.o.g $ i_p > i_q $. 
For an articulation node $u$, $i_p$ and $i_q$ are the indices corresponding to the stub lengths induced by the intersection points of $s(a(u))$ and its neighbors in $C$ among the vertex set of $p(u)$.
For a block node $u$, $i_p$ and $i_q$ are the indices corresponding to the intersection points of the parent's segment $s(a(p(u)))$ with its neighbors in $V(u)$.
%These are exactly the indices of the stub-lengths leading to an intersection between two cycles of segments in $ \Gamma $. 
Similar to Section~\ref{sec:tree} we now define three values for each articulation node $ u \in \mathcal V $ with degree $ \delta $:
\begin{align*}
	\sollong(u) = \max\{T_0(u),\dots,T_\delta(u)\}\\
	\solmid(u) = \max\{T_1(u),\dots,T_{i_q}(u)\}\\
	\solshort(u) = \max\{T_1(u),\dots,T_{i_p}(u)\}
\end{align*}
Further we define the recurrence $T_i(u)$ for all articulation nodes $u \in \mathcal V$:
\begin{align}
\label{rec:cactus_articulation}
T_i(u) = \ell_i(a(u)) + \sum_{v \in c(u)}
\begin{cases}
T_{i_p(v)}(v) & \text{if } i < i_p(v)\\ %\ell_i(a(u)) \leq a \\
T_{i_q(v)}(v) & \text{if } i_p(v) < i \leq i_q(v) \\
T_{0}(v) & \text{otherwise}
% \solshort(v) & \text{if } i < i_p(v)\\ %\ell_i(a(u)) \leq a \\
% \solmid(v) & \text{if } i_p(v) < i \leq i_q(v) \\
% \sollong(v) & \text{otherwise}
\end{cases}.
\end{align}
For a block node $ u \in \mathcal V $ we cut open the cycle it represents at vertex $a(p(u))$ (refer to Bruckdorfer et al.~\cite{bcgkmn-pped-17} for details) and consider the three relevant stub lengths of the parent's segment $s(a(p(u)))$ induced by $i_p(u)$ and $i_q(u)$ and the whole segment. 
The resulting subgraph is a path on $V(u)$ with attached children that are articulation vertices in $\mathcal T$ so we can apply our algorithm of Section~\ref{sec:tree} with minor modifications.
Once all values in the subgraph of block node $u$ are computed, we can derive the values $T_i(u)$ for the block node $u$.

@@

\begin{theorem}
	Let $ G $ be a simple graph with $m$ edges and $ \Gamma $ a straight-line drawing of $ G $. If the intersection graph $ C = (V,E) $ of $ G $ is a cactus with maximum degree $ k \in \mathbb{N} $, problem \maxsped can be solved in $ O(mk + m \log m)) $ time and space.
\end{theorem}
%\begin{proof}
%	Follows as in Theorem~\ref{thm:tree} by maximizing over the tree using Recurrence~\ref{rec:cactus_articulation}. Then the running time follows from constructing the intersection graph and the block-cut tree as well as Lemma~\ref{lem:compute_block} for the block vertices and applying Lemma~\ref{lem:compute_vertex} for the articulation vertices.
%\end{proof}

\bibliography{paper}
\end{document}

>>>>

\begin{theorem}
	\label{thm:tree}
	Let $ G $ be a simple graph with $m$ edges and $ \Gamma $ a straight-line drawing of $ G $. If the intersection graph $ C = (V,E) $ of $ G $ is a tree with maximum degree $ k \in \mathbb{N} $, problem \maxsped can be solved in $ O(mk + m \log m) $ time and space.
\end{theorem}
% \begin{proof}
% 	Since $ C $ is known to be a tree, and hence $ |E| = O(m) $, we can compute the intersection graph and the stub lengths in $ O(m\log m) $ time and space.
% 	With Lemma~\ref{lem:compute_vertex} and Recurrence~\ref{rec:tree} we need one pass through the tree to find the roots maximum in $ O(mk) $ time. Adding the bounds yields the desired runtime of $ O(mk + m \log m)) $.
% %	To solve the \maxsped problem on $ G $ we need to compute $ \max \{T_i(r) \mid 0 \leq i \leq k \} $ for the root $ r \in V $ of $ C $ as seen above. Doing this bottom up requires one pass through $ C $ hence by applying Lemma~\ref{lem:compute_vertex} we need $ O(mk) $ time to compute all $ T_i(u) $ for all $ u \in V $. For every vertex we only need to compute $ O(k) $ values and store two of them, namely $ \sollong(u) $ and $ \solshort(u) $, so we do not need more than $ O(mk) $ memory. Adding the bounds yields the desired runtime of $ O(mk + m \log m)) $.
% %
%
%
%
% 	Retrieving the actual stub length for each edge in $ G $ needed to produce the SPED can be done by backtracking the subsolutions which requires at most $ O(mk) $ time.
% \end{proof}

@@

\subsection{Cactus graphs}
\label{sec:cactus}

%In this section we restrict $ C = (V,E) $ to be a cactus graph, i.e. a simple graph in which no edge $ e \in E $ is part of more than one cycle. For a cactus graph $ C $ the \emph{block-cut tree} is the tree $ \mathcal T = (\mathcal V, \mathcal E) $ where $ \mathcal V $ contains a vertex for every cycle and cut vertex in $ C $. We call a vertex $ u \in \mathcal V $ a \emph{block vertex} if it represents an induced cycle with vertices $ V_c \subseteq V $ and \emph{articulation vertex} representing the cut vertex $ v \in V $ else. For a block vertex $ u \in \mathcal V $ we define $ \textit{cycle}(u) = V_c$ and for an articulation vertex $ u \in \mathcal V $ we define $ A(u) = v $. There is an edge $ uv \in \mathcal E $ between a block vertex $ u $ and articulation vertex $ v $ whenever $ v \in \textit{cycle}(u) $ and an edge $ uv \in \mathcal E $ between two articulation vertices whenever $ A(u)A(v) $ was an edge in $ E $.
%
%Let $ C = (V,E) $ be a cactus graph and moreover the intersection graph of a graph $ G $ with straight-line drawing $ \Gamma $. Additionally let $ \mathcal T = (\mathcal V, \mathcal E) $ be the block-cut tree of $ C $ we define for each vertex $ u \in \mathcal V $ two indices $ i_p,i_q $, w.l.o.g $ i_p > i_q $. This are exactly the indices of the stub-lengths leading to an intersection between two cycles of segments in $ \Gamma $. Similar to Section~\ref{sec:tree} we define three values for every vertex $ u \in \mathcal V $ with degree $ \delta $:
%\begin{align*}
%	\solshort(u) &= \max\{T_0(u),\dots,T_\delta(u)\}\\
%	\solmid(u) &= \max\{T_1(u),\dots,T_{i_p}(u)\}\\
%	\sollong(u) &= \max\{T_1(u),\dots,T_{i_q}(u)\}
%\end{align*}
%%\begin{eqnarray*}
%%	\solshort(u) = \max\{T_0(u),\dots,T_\delta(u)\}\qquad
%%	\solmid(u) = \max\{T_1(u),\dots,T_{i_p}(u)\}\\
%%	\multicolumn{2}{c}{$\sollong(u) = \max\{T_1(u),\dots,T_{i_q}(u)\}$}
%%\end{eqnarray*}
%For an articulation vertex $ u \in \mathcal V $ we can define a similar recurrence to Recurrence~\ref{rec:tree}:

We now generalize our previous algorithm to cactus graphs. 
Due to space constraints we provide only a sketch of the arguments and omit further details.
Let the intersection graph $ C = (V,E) $ be a \emph{cactus graph}, i.e. a simple graph in which no edge $ e \in E $ is part of more than one cycle. 
For a cactus graph $ C $ the \emph{block-cut tree} is an arbitrarily rooted tree $ \mathcal T = (\mathcal V, \mathcal E) $, where $ \mathcal V $ contains a node for every cycle and for every cut vertex in $ C $. 
We call a node $ u \in \mathcal V $ a \emph{block node} if it represents an induced cycle on vertices $ V(u) \subseteq V $ and an \emph{articulation node} of the cut vertex $ a(u) = v \in V $ otherwise. 
%For a block node $ u \in \mathcal V $ we define $ \textit{cycle}(u) = V_c$ and f
%For an articulation node $ u \in \mathcal V $ we define $ a(u) = v $ to be the corresponding cut vertex. 
We have an edge $ uv \in \mathcal E $ between a block node $ u $ and an articulation node $ v $ whenever $ a(v) \in V(u) $ and an edge $ uv \in \mathcal E $ between two articulation nodes whenever $ a(u)a(v) \in E$. % is an edge in $ E $.
We re-use the notation $p(u)$ and $c(u)$ to refer to the parent and to the children of a node $u \in \mathcal V$.

%Let $ C = (V,E) $ be a cactus graph and moreover the intersection graph of a graph $ G $ and straight-line drawing $ \Gamma $. Additionally let $ \mathcal T = (\mathcal V, \mathcal E) $ be the block-cut tree of $ C $ 
We further define for each node $ u \in \mathcal V $ two indices $ i_p(u) \le i_q(u) $. %, w.l.o.g $ i_p > i_q $. 
For an articulation node $u$, $i_p$ and $i_q$ are the indices corresponding to the stub lengths induced by the intersection points of $s(a(u))$ and its neighbors in $C$ among the vertex set of $p(u)$.
For a block node $u$, $i_p$ and $i_q$ are the indices corresponding to the intersection points of the parent's segment $s(a(p(u)))$ with its neighbors in $V(u)$.
%These are exactly the indices of the stub-lengths leading to an intersection between two cycles of segments in $ \Gamma $. 
Similar to Section~\ref{sec:tree} we now define three values for each articulation node $ u \in \mathcal V $ with degree $ \delta $:
\begin{align*}
	\sollong(u) = \max\{T_0(u),\dots,T_\delta(u)\}\\
	\solmid(u) = \max\{T_1(u),\dots,T_{i_q}(u)\}\\
	\solshort(u) = \max\{T_1(u),\dots,T_{i_p}(u)\}
\end{align*}
Further we define the recurrence $T_i(u)$ for all articulation nodes $u \in \mathcal V$:

\begin{align}
\label{rec:cactus_articulation}
T_i(u) = \ell_i(a(u)) + \sum_{v \in c(u)}
\begin{cases}
%\solshort(v) & \text{if } \ell_i(A(u)) \leq a \\
%\solmid(v) & \text{if } a < \ell_i(A(u)) \leq b \\
%\sollong(v) & \text{else}
%\end{cases}.
%\end{align}
%For a block vertex $ u \in \mathcal V $ we have to cut open the cycle it represents (compare \cite{bcgkmn-pped-17}). We do this at $ A(p(u)) $. The resulting graph is a path between the vertices in $ \textit{cycle}(u) $ with attached articulation vertices. This is especially a tree and we use the aglorithm from Section~\ref{sec:tree}, maximizing over three values of the articulation vertices though,  to compute the $ T_i(u) $.
%
%If the values computed for the block vertices is correct, we can follow the correctness of the algorithm as in Section~\ref{sec:tree}. Let $ u \in \mathcal V $ be a block vertex then by \cite{bcgkmn-pped-17} we know the optimal solution for \maxsped just on $ C[\textit{cycle}(u)] $ is correctly computed by the above procedure since the algorithm from Section~\ref{sec:tree} correctly handles the created path. Now if $ u $ has articulation vertices attached to it the above computes the correct value since this is simply a tree. Further there is no $ v \in c(u) $ which is not an articulation vertex, hence the maximum ink value for $ u $ is correctly computed if we know the values for the children in $ c(u) $.
%
%For the runtime we can again argue the same as in Section~\ref{sec:tree}, but this time we have to also check cycles of length $ n_c $ which leads to a runtime of $ O(n_ck^2) $ for a block vertex $ u \in \mathcal V $ with $ n_c = |\textit{cycle}(u)| $. In case $ u $ is an articulation vertex we can solve it in $ O(k) $. As before the optimal solution to the \maxsped problem on $ G $ with drawing $ \Gamma $ can be found by backtracking from the root. Additionally the intersection graph can be computed in $ O(m\log m) $ where $ m $ is the number of edges in the original graph $ G $ since $ |E| = O(m) $ in a cactus and we obtain the following theorem.
T_{i_p(v)}(v) & \text{if } i < i_p(v)\\ %\ell_i(a(u)) \leq a \\
T_{i_q(v)}(v) & \text{if } i_p(v) < i \leq i_q(v) \\
T_{0}(v) & \text{otherwise}
% \solshort(v) & \text{if } i < i_p(v)\\ %\ell_i(a(u)) \leq a \\
% \solmid(v) & \text{if } i_p(v) < i \leq i_q(v) \\
% \sollong(v) & \text{otherwise}
\end{cases}.
\end{align}
For a block node $ u \in \mathcal V $ we cut open the cycle it represents at vertex $a(p(u))$ (refer to Bruckdorfer et al.~\cite{bcgkmn-pped-17} for details) and consider the three relevant stub lengths of the parent's segment $s(a(p(u)))$ induced by $i_p(u)$ and $i_q(u)$ and the whole segment. 
The resulting subgraph is a path on $V(u)$ with attached children that are articulation vertices in $\mathcal T$ so we can apply our algorithm of Section~\ref{sec:tree} with minor modifications.
Once all values in the subgraph of block node $u$ are computed, we can derive the values $T_i(u)$ for the block node $u$.

@@

\begin{theorem}
	Let $ G $ be a simple graph with $m$ edges and $ \Gamma $ a straight-line drawing of $ G $. If the intersection graph $ C = (V,E) $ of $ G $ is a cactus with maximum degree $ k \in \mathbb{N} $, problem \maxsped can be solved in $ O(mk + m \log m) $ time and space.
\end{theorem}
%\begin{proof}
%	Follows as in Theorem~\ref{thm:tree} by maximizing over the tree using Recurrence~\ref{rec:cactus_articulation}. Then the running time follows from constructing the intersection graph and the block-cut tree as well as Lemma~\ref{lem:compute_block} for the block vertices and applying Lemma~\ref{lem:compute_vertex} for the articulation vertices.
%\end{proof}

%\bibliography{paper}
\begin{thebibliography}{1}

	\bibitem{bew-vnd-95}
	Richard~A. Becker, Stephen~G. Eick, and Allan~R. Wilks.
	\newblock Visualizing network data.
	\newblock {\em {IEEE} Trans. Visualization and Computer Graphics}, 1(1):16--28,
	1995.

	\bibitem{blmt-pedhmitc-16}
	Carla Binucci, Giuseppe Liotta, Fabrizio Montecchiani, and Alessandra Tappini.
	\newblock Partial edge drawing: Homogeneity is more important than crossings
	and ink.
	\newblock In {\em Information, Intelligence, Systems Applications (IISA'16)},
	pages 1--6. IEEE, 2016.

	\bibitem{b-sgh-15}
	Till Bruckdorfer.
	\newblock {\em Schematics of Graphs and Hypergraphs}.
	\newblock PhD thesis, Uni Tübingen, 2015.

	\bibitem{bcgkmn-pped-17}
	Till Bruckdorfer, Sabine Cornelsen, Carsten Gutwenger, Michael Kaufmann,
	Fabrizio Montecchiani, Martin Nöllenburg, and Alexander Wolff.
	\newblock Progress on partial edge drawings.
	\newblock {\em J. Graph Algorithms Appl.}, 21(4):757--786, 2017.

	\bibitem{bk-ecbe-12}
	Till Bruckdorfer and Michael Kaufmann.
	\newblock Mad at edge crossings? {Break} the edges!
	\newblock In {\em Fun with Algorithms (FUN'12)}, volume 7288 of {\em LNCS},
	pages 40--50. Springer, 2012.

	\bibitem{bkl-us-15}
	Till Bruckdorfer, Michael Kaufmann, and Simon Leibßle.
	\newblock Ped user study.
	\newblock In {\em Graph Drawing (GD'15)}, volume 9411 of {\em LNCS}, pages
	551--553. Springer International Publishing, 2015.

	\bibitem{bvkw-epdldge-12}
	Michael Burch, Corinna Vehlow, Natalia Konevtsova, and Daniel Weiskopf.
	\newblock Evaluating partially drawn links for directed graph edges.
	\newblock In {\em Graph Drawing (GD'11)}, volume 7034 of {\em LNCS}, pages
	226--237. Springer, 2012.

	\bibitem{bcko-cgaa-08}
	Mark de~Berg, Otfried Cheong, Marc van Kreveld, and Mark Overmars.
	\newblock {\em Computational Geometry: Algorithms and Applications}.
	\newblock Springer, 3rd edition edition, 2008.

	\bibitem{l-pftu-82}
	David Lichtenstein.
	\newblock Planar formulae and their uses.
	\newblock {\em SIAM J Comput}, 11(2):329--343, 1982.	
\end{thebibliography}
\end{document}
